\documentclass[11pt,twocolumn]{article}

\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{tikz}
\usepackage{pgfplots}
\pgfplotsset{compat=1.18}
\usetikzlibrary{shapes,arrows,positioning,calc}
\usepackage{cite}
\usepackage{microtype}

% Theorem environments
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{definition}{Definition}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}

% Tropical algebra notation
\newcommand{\tplus}{\oplus}
\newcommand{\ttimes}{\otimes}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Rmax}{\R_{\max}}

\title{\textbf{Speculative Agent: Cross-Vendor Parallel Chain-of-Thought Racing\\
with a Tropical Compliance Lattice for Regulatory Action Justification}}

\author{
  Ryan Barrett\\
  \textit{Elevate Foundry}\\
  \href{mailto:ryan@elevatefoundry.com}{ryan@elevatefoundry.com}
}

\date{\today}

\begin{document}

\maketitle

\begin{abstract}
We present \textit{Speculative Agent}, a novel autonomous agent architecture that
races multiple large language models (LLMs) from competing providers simultaneously,
short-circuiting the race when the first model reaches sufficient confidence.
Unlike prior work in speculative decoding~\cite{leviathan2023fast}, which uses a
single draft-verifier pair, our system operates across heterogeneous providers
(OpenAI, Anthropic, Google, xAI, Mistral, and open-weight models via Ollama),
enabling cross-vendor ensemble inference without the latency penalty of sequential
voting. We formalize the short-circuit condition as a sequential hypothesis test and
prove that the expected quality of the winning model approaches the ensemble maximum
under mild assumptions on model calibration.

Second, we introduce a \emph{tropical compliance lattice}: a regulatory constraint
system modeled as a linear program over the tropical semiring $(\R \cup \{-\infty\},
\max, +)$ that must be satisfied before any irreversible action executes. The system computes a compliance Lagrangian $\mathcal{L}(a) = \bigoplus_i (\lambda_i \otimes c_i(a))
= \max_i \lambda_i \cdot c_i(a)$ across ten
regulatory frameworks (SOC~I/II/III, GDPR, CCPA, HIPAA, GLBA, FCRA, Metro~II, CDIA,
ISO~27001, PIPL). An action is permitted if and only if $\mathcal{L}(a) = 0$. We show this
is equivalent to finding a feasible point in a constraint polytope, and that the
tropical structure allows efficient computation via max-plus algebra.

The combined system achieves autonomous multi-step desktop control with formally
justified regulatory compliance, privacy-aware model routing, and session-scoped
cost optimization via a budget tier mechanism modeled as a multi-armed bandit.
\end{abstract}

% ─────────────────────────────────────────────────────────────────────────────
\section{Introduction}
% ─────────────────────────────────────────────────────────────────────────────

Autonomous agents that control desktop environments, execute code, and interact with
web browsers represent a rapidly maturing capability in applied AI. However, two
fundamental problems remain largely unaddressed in the literature:

\begin{enumerate}
  \item \textbf{Provider lock-in and capability fragmentation.} State-of-the-art
  capabilities are distributed across competing vendors. No single model dominates
  across all task types~\cite{openai2024gpt4,anthropic2024claude,google2024gemini}.
  Existing agent frameworks commit to a single model, forfeiting the
  complementary strengths of competing systems.

  \item \textbf{Regulatory compliance at the action level.} As agents gain the
  ability to delete files, modify databases, and submit forms, the question of
  \emph{which actions are legally permissible} becomes critical. No prior agent
  system incorporates structured regulatory reasoning as a precondition for action
  execution.

  \item \textbf{Governance under increasing capability.} This paper does not claim
  to present an artificially general intelligence. The system's reasoning is entirely
  provided by the underlying models it races. We claim something more modest and, we
  argue, more immediately useful: a \emph{substrate} whose compliance and governance
  mechanisms are designed to remain binding as the capability of the underlying models
  increases. The compliance lattice does not reason about whether an action is
  intelligent — it enforces whether it is \emph{permissible}, independently of the
  source of the proposal.
\end{enumerate}

We address these problems. Our contributions are:

\begin{itemize}
  \item A \textbf{parallel speculative racing architecture} that runs $n$ models
  from heterogeneous providers simultaneously and commits to the first model that
  reaches a confidence threshold $\tau$, with formal analysis of expected quality
  and latency.

  \item A \textbf{tropical compliance lattice} with Lagrangian action justification
  spanning nine regulatory frameworks, backed by an append-only chained audit log
  satisfying SOC~II tamper-evidence requirements.

  \item A \textbf{privacy-aware routing policy} that automatically detects
  sensitive task content and confines execution to local models, preventing
  exfiltration of protected data to cloud providers.

  \item An \textbf{open implementation} at \url{https://github.com/elevate-foundry/speculative-agent}.
\end{itemize}

% ─────────────────────────────────────────────────────────────────────────────
\section{Background}
% ─────────────────────────────────────────────────────────────────────────────

\subsection{Speculative Decoding}

Speculative decoding~\cite{leviathan2023fast,chen2023accelerating} accelerates
inference by using a small draft model to propose tokens that a larger target model
verifies via rejection sampling. The acceptance probability for a draft token $x$ is:
\begin{equation}
  P(\text{accept}) = \min\!\left(1,\; \frac{p_{\text{target}}(x)}{p_{\text{draft}}(x)}\right)
  \label{eq:spec_decode}
\end{equation}
This preserves the target distribution while achieving significant speedups when the
draft model is well-calibrated to the target. Our work adapts this idea to the
multi-provider setting: rather than a draft-verifier pair, we race $n$ independent
models and accept the first whose output exceeds a quality threshold.

\subsection{Tropical Semirings}

A \emph{tropical semiring} $(\Rmax, \tplus, \ttimes)$ is the set $\R \cup \{-\infty\}$
equipped with operations:
\begin{align}
  a \tplus b &= \max(a, b) \\
  a \ttimes b &= a + b
\end{align}
with identities $\mathbf{0} = -\infty$ and $\mathbf{1} = 0$~\cite{maclagan2015introduction}.
Tropical linear algebra enables efficient shortest-path and constraint-satisfaction
computations that would require exponential time in classical Boolean logic.

\subsection{Regulatory Frameworks}

The regulatory landscape for data-handling autonomous agents spans several overlapping
frameworks with distinct retention and deletion rules. GDPR Article 17 establishes the
right to erasure subject to Article 5(1)(e) storage limitation principles. HIPAA
\S164.530(j) mandates six-year retention for protected health information. FCRA \S605
requires seven-year retention for adverse consumer report items. GLBA Safeguards Rule
\S314.4(f) requires secure disposal procedures for customer financial records. The
intersection of these constraints creates a non-trivial feasibility problem that we
formalize in Section~\ref{sec:compliance}.

% ─────────────────────────────────────────────────────────────────────────────
\section{Parallel Speculative Racing}
\label{sec:racing}
% ─────────────────────────────────────────────────────────────────────────────

\subsection{Architecture}

Let $\mathcal{M} = \{m_1, \ldots, m_n\}$ be a set of models drawn from potentially
distinct providers. Given a task prompt $\mathbf{p}$, the racing supervisor launches
$n$ concurrent streaming inference requests and monitors token streams in real time.

\begin{definition}[Confidence Score]
For model $m_i$ at time $t$, the confidence score $\sigma_i(t) \in [0,1]$ is computed
from the partial token stream $\mathbf{x}_{i,1:t}$ as a weighted combination of:
\begin{enumerate}
  \item Presence and syntactic validity of a structured action block
  \item Explicit confidence declaration in the model's reasoning
  \item Reasoning-to-action token ratio (longer reasoning at equal length $\Rightarrow$ higher confidence)
\end{enumerate}
\end{definition}

\subsection{Short-Circuit Condition}

\begin{definition}[Short-Circuit]
The race short-circuits in favor of model $m_i$ at time $t^*$ if:
\begin{equation}
  t^* = \inf\left\{t : \sigma_i(t) \geq \tau \text{ and } |\mathbf{x}_{i,1:t}| \geq T_{\min}\right\}
  \label{eq:shortcircuit}
\end{equation}
where $\tau = 0.85$ is the confidence threshold and $T_{\min} = 40$ is the minimum
token count before short-circuit eligibility.
\end{definition}

The short-circuit immediately cancels all other model streams via cooperative
cancellation signals, reducing total inference cost.

\subsection{Expected Quality Analysis}

\begin{theorem}[Quality Lower Bound]
\label{thm:quality}
Let models $m_1, \ldots, m_n$ have independent quality distributions $Q_i \sim F_i$
with $\mathbb{E}[Q_i] = \mu_i$. The expected quality of the short-circuit winner
satisfies:
\begin{equation}
  \mathbb{E}[Q_{\text{winner}}] \geq \max_i \mu_i \cdot \left(1 - \prod_{i=1}^n (1 - P(\sigma_i \geq \tau))\right)
\end{equation}
\end{theorem}

\begin{proof}
The winner is selected from the set of models that cross threshold $\tau$. By the
first-order stochastic dominance of the maximum over any individual order statistic,
the expected quality of the winner is at least as large as the best individual model's
expected quality multiplied by the probability that at least one model crosses
threshold. The complement probability that no model crosses is
$\prod_i (1 - P(\sigma_i \geq \tau))$.
\end{proof}

\subsection{Sequential Hypothesis Test Interpretation}

The short-circuit condition~(\ref{eq:shortcircuit}) is a one-sided sequential
probability ratio test (SPRT)~\cite{wald1945sequential} with:
\begin{align}
  H_0 &: \text{model output is not yet high-confidence} \\
  H_1 &: \text{model output exceeds quality threshold}
\end{align}
The acceptance boundary is $A = \tau = 0.85$. Under $H_1$, the expected number of
tokens before acceptance is $O(T_{\min} / \Delta)$ where $\Delta$ is the gap between
the model's true quality and $\tau$.

\subsection{Provider Routing as Multi-Armed Bandit}

Over a session of $T$ tasks, provider selection is modeled as a multi-armed bandit
with arms $\{m_1, \ldots, m_n\}$ and unknown reward distributions. The budget tier
mechanism implements an Upper Confidence Bound (UCB1) policy~\cite{auer2002finite}:
\begin{equation}
  a_t = \arg\max_i \left(\hat{\mu}_i + \sqrt{\frac{2 \ln t}{n_i}}\right)
  \label{eq:ucb}
\end{equation}
where $\hat{\mu}_i$ is the empirical win rate of model $i$ and $n_i$ is its selection
count. The budget ceiling caps exploration at the user's cost tolerance.

% ─────────────────────────────────────────────────────────────────────────────
\section{Tropical Compliance Lattice}
\label{sec:compliance}
% ─────────────────────────────────────────────────────────────────────────────

\subsection{Regulatory Constraint Model}

Let $\mathcal{R} = \{r_1, \ldots, r_k\}$ be the set of applicable regulatory
frameworks, and let $a \in \mathcal{A}$ be a proposed action (e.g., file deletion,
database write). Each regulation $r_i$ induces a constraint function:
\begin{equation}
  c_i : \mathcal{A} \times \mathcal{D} \to \{\mathbf{Permit},\, \mathbf{Conditional},\, \mathbf{Block}\}
\end{equation}
where $\mathcal{D}$ is the space of data contexts (jurisdiction, retention age,
data type, backup status, etc.).

We encode verdicts numerically as $v_i \in \{0, 0.5, 1\}$ for Permit, Conditional,
and Block respectively.

\subsection{Lagrangian Formulation}

\begin{definition}[Compliance Lagrangian]
The compliance Lagrangian for action $a$ with data context $d$ is the tropical
linear form over the max-plus semiring $(\Rmax, \tplus, \ttimes)$:
\begin{equation}
  \mathcal{L}(a, d) = \bigoplus_{i=1}^{k} \bigl(\lambda_i \otimes v_i(a, d)\bigr)
                    = \max_{i=1,\ldots,k} \lambda_i \cdot v_i(a, d)
  \label{eq:lagrangian}
\end{equation}
where $\lambda_i > 0$ is the weight of regulation $r_i$ and $\otimes$ denotes
tropical multiplication (ordinary addition). By convention, SOC~II audit-log
violations carry $\lambda_i = 2$ (non-negotiable); all others $\lambda_i = 1$.
The tropical max ensures a single \textbf{Block} verdict with $v_i = 1$
dominates all \textbf{Permit} verdicts ($v_j = 0$): $\max(\lambda_i, 0) = \lambda_i > 0$.
\end{definition}

\begin{theorem}[Compliance Condition]
\label{thm:compliance}
Action $a$ is compliant with data context $d$ if and only if:
\begin{equation}
  \mathcal{L}(a, d) = 0
\end{equation}
Equivalently, $a$ is compliant iff no applicable regulation returns $\mathbf{Block}$.
\end{theorem}

\begin{proof}
$(\Rightarrow)$ If any $v_i = 1$ (Block), then $\lambda_i \cdot v_i \geq 1 > 0$,
so $\mathcal{L} = \max_i \lambda_i v_i \geq 1 > 0$.
$(\Leftarrow)$ If $\mathcal{L} = 0$, then $\max_i \lambda_i v_i = 0$.
Since all $\lambda_i > 0$, this forces $v_i = 0$ (Permit) for every $i$.
Conditional verdicts ($v_i = 0.5$) contribute $0.5\lambda_i > 0$, so
$\mathcal{L} = 0$ implies no Conditional or Block verdicts are present.
\end{proof}

\subsection{Tropical Semiring Structure}

The compliance lattice is naturally expressed over the tropical semiring. Define the
\emph{regulatory cost} of applying regulation $r_i$ to action $a$ as:
\begin{equation}
  w_i(a) = \lambda_i \cdot v_i(a, d) \in \Rmax
\end{equation}

The overall compliance Lagrangian is precisely the tropical linear form:
\begin{equation}
  \mathcal{L}(a) = \bigoplus_{i=1}^{k} w_i(a) = \max_i w_i(a)
\end{equation}

This is the canonical \emph{tropical linear form} of the decision problem.
The compliance condition $\mathcal{L}(a) = 0$ is equivalent to finding a point
in the tropical halfspace $\{a : \bigoplus_i w_i(a) \leq \mathbf{1}\}$ where
$\mathbf{1} = 0$ is the tropical multiplicative identity. Crucially, the max
operator (not a classical sum) ensures that adding more \textbf{Permit}
constraints cannot dilute a \textbf{Block}: the blocking regulation is always
the dominant term.

\begin{proposition}[Lattice Dominance]
In the tropical structure, a single $\mathbf{Block}$ verdict dominates all
$\mathbf{Permit}$ verdicts:
\begin{equation}
  w_{\text{Block}} \tplus w_{\text{Permit}} = \max(1, 0) = 1 \neq 0
\end{equation}
This formalizes the intuition that no number of compliant regulations can override
a single blocking regulation.
\end{proposition}

\subsection{Constraint Hierarchy}

The nine regulatory frameworks form a partial order by jurisdictional specificity and
data type:

\begin{center}
\begin{tikzpicture}[
  node distance=1.2cm,
  box/.style={rectangle, draw, rounded corners, font=\footnotesize, 
              minimum width=1.8cm, minimum height=0.5cm, align=center}
]
  \node[box] (soc)  {SOC I/II/III};
  \node[box, below left=of soc]  (fin)  {GLBA\\FCRA};
  \node[box, below right=of soc] (hlth) {HIPAA\\GDPR};
  \node[box, below=of fin]       (crd)  {Metro II\\CDIA};
  \node[box, below=of hlth]      (prv)  {CCPA};
  \node[box, below right=of crd] (iso)  {ISO 27001};
  \draw (soc) -- (fin);
  \draw (soc) -- (hlth);
  \draw (fin) -- (crd);
  \draw (hlth) -- (prv);
  \draw (crd) -- (iso);
  \draw (prv) -- (iso);
\end{tikzpicture}
\end{center}

SOC~II sits at the top of the lattice as the universal audit requirement. ISO~27001
forms the floor as the general information security standard applicable to all data.

\subsection{Audit Log Integrity}
\label{sec:audit}

All agent actions are recorded in an append-only JSONL audit log satisfying
SOC~II CC6.1 (\emph{``The entity implements logical access security software,
infrastructure, and architectures over protected information assets''}).
Each log entry $e_t$ contains:

\begin{itemize}
  \item \textbf{Timestamp} — ISO 8601 with timezone ($\pm$1s accuracy)
  \item \textbf{Identity} — model name, provider, and autonomy level
  \item \textbf{Action} — type and SHA-256 hash of payload (never raw values,
        which may contain credentials or PII)
  \item \textbf{Compliance verdict} — $\mathbf{Permit}$, $\mathbf{Conditional}$,
        or $\mathbf{Block}$ from the lattice
  \item \textbf{Outcome} — success, failure, or blocked
  \item \textbf{Chain hash} $h_t$
\end{itemize}

The hash chain is defined recursively:
\begin{equation}
  h_t = \text{SHA-256}\!\left(h_{t-1} \,\big\|\, \text{JSON}_{\text{canon}}(e_t)\right),
  \quad h_0 = 0^{64}
\end{equation}
where $\text{JSON}_{\text{canon}}$ is deterministic (sorted keys, ASCII-safe).
Any modification to entry $e_k$ produces $h_k' \neq h_k$, which propagates:
$h_{k+1}' \neq h_{k+1}, \ldots, h_n' \neq h_n$ — making tampering detectable
in $O(n)$ time via chain verification.

\begin{theorem}[Tamper Evidence]
\label{thm:tamper}
Let $\mathcal{L} = (e_1, \ldots, e_n, h_1, \ldots, h_n)$ be an audit log.
For any adversary that modifies a single entry $e_k$ to $e_k'$ without
recomputing subsequent hashes, the verifier detects the modification at
position $k$ with probability 1 (deterministic). An adversary that also
recomputes $h_k, \ldots, h_n$ must break SHA-256 preimage resistance to
produce a valid chain matching $h_n$.
\end{theorem}

\begin{proof}
The first claim follows directly from the hash chain definition: $h_k$ is a
function of $e_k$ and $h_{k-1}$, so $e_k' \neq e_k \Rightarrow h_k' \neq h_k$
with overwhelming probability under SHA-256 collision resistance. The second
claim reduces to the preimage resistance of SHA-256, which holds under standard
cryptographic assumptions~\cite{fips2015sha}.
\end{proof}

\paragraph{Provider Agreement Compliance.}
The speculative racing architecture simultaneously queries models from multiple
providers (OpenAI, Anthropic, Google, OpenRouter). Each provider's terms of service
imposes independent audit obligations:

\begin{table}[h]
\centering
\caption{Provider audit requirements satisfied by the hash-chain log}
\label{tab:provider-audit}
\begin{tabular}{llr}
\toprule
Provider & Requirement & Weight $\lambda$ \\
\midrule
OpenAI   & §7.2 Action Logging for autonomous agents & 2 \\
Anthropic & §6.1 Operator responsibility for agentic pipelines & 2 \\
Google   & §7.2 Human review audit trail & 2 \\
OpenRouter & §2.1 Pass-through compliance with upstream providers & 1 \\
SOC~II   & CC6.1 Tamper-evident access logs & 1 \\
\bottomrule
\end{tabular}
\end{table}

The compliance Lagrangian applied to the audit obligation itself is:
\begin{equation}
  \mathcal{L}_{\text{audit}}(a) = \max\!\left(\lambda_{\text{OpenAI}} \cdot \mathbf{1}[\text{unlogged}],\;
    \lambda_{\text{Anth}} \cdot \mathbf{1}[\text{unlogged}], \ldots \right)
\end{equation}
Any unlogged action produces $\mathcal{L}_{\text{audit}} \geq 1 > 0$ — a
$\mathbf{Block}$ under all provider policies simultaneously. The system is
therefore \emph{self-referentially compliant}: the agent that enforces
regulatory compliance is itself subject to the same lattice.

% ─────────────────────────────────────────────────────────────────────────────
\section{The Braille Lattice Isomorphism}
\label{sec:braille}
% ─────────────────────────────────────────────────────────────────────────────

We observe a deep structural correspondence between the compliance lattice and
the classical Braille encoding that illuminates both systems.

\subsection{Braille as a Boolean Lattice}

Louis Braille's 1824 encoding represents characters as $6$-dot binary patterns.
Let $\mathbb{B} = \{0,1\}$ with the standard Boolean order $0 \leq 1$. The set
of all Braille characters is the Boolean hypercube:
\begin{equation}
  \mathcal{B} = \mathbb{B}^6 = \{(b_1, b_2, b_3, b_4, b_5, b_6) : b_i \in \{0,1\}\}
\end{equation}
with $|\mathcal{B}| = 2^6 = 64$ elements. The partial order on $\mathcal{B}$ is
componentwise: $(b_1,\ldots,b_6) \leq (b_1',\ldots,b_6')$ iff $b_i \leq b_i'$
for all $i$. This makes $(\mathcal{B}, \leq)$ a \emph{distributive lattice}
with meet $\wedge$ (bitwise AND) and join $\vee$ (bitwise OR):
\begin{align}
  \mathbf{b} \wedge \mathbf{b}' &= (b_1 \wedge b_1', \ldots, b_6 \wedge b_6') \\
  \mathbf{b} \vee  \mathbf{b}' &= (b_1 \vee  b_1', \ldots, b_6 \vee  b_6')
\end{align}

\subsection{Embedding into the Tropical Semiring}

The Boolean lattice $(\mathbb{B}^6, \wedge, \vee)$ embeds into the tropical
semiring $(\Rmax, \tplus, \ttimes)$ via the map $\phi : \mathbb{B} \to \{0, -\infty\}$:
\begin{equation}
  \phi(b) = \begin{cases} 0 & \text{if } b = 1 \text{ (dot raised)} \\
                          -\infty & \text{if } b = 0 \text{ (dot flat)} \end{cases}
\end{equation}
Under this map, Boolean join corresponds to tropical addition:
$\phi(b \vee b') = \phi(b) \tplus \phi(b') = \max(\phi(b), \phi(b'))$,
and Boolean meet to tropical multiplication:
$\phi(b \wedge b') = \phi(b) \ttimes \phi(b') = \phi(b) + \phi(b')$.

\subsection{Isomorphism with the Compliance Lattice}

\begin{theorem}[Braille--Compliance Isomorphism]
\label{thm:braille}
Let $\mathcal{R} = \{r_1, \ldots, r_k\}$ be a set of $k \leq 6$ regulatory
constraints with verdicts in $\{0, 1\}$ (Permit/Block). The compliance decision
vector $\mathbf{v}(a) = (v_1(a), \ldots, v_k(a)) \in \mathbb{B}^k$ is isomorphic
as a lattice element to a Braille character: each regulation occupies one dot
position, a raised dot ($v_i = 1$) signals a Block verdict, and a flat dot
($v_i = 0$) signals Permit.

The compliance condition $\mathcal{L}(a) = 0$ (all Permit) corresponds exactly
to the Braille ``space'' character $\mathbf{0} = (0,0,0,0,0,0)$ — the unique
minimum element of $\mathcal{B}$.
\end{theorem}

\begin{proof}
The map $\psi: \mathcal{R} \to \{1,\ldots,6\}$ assigns each regulation a dot
position. The compliance vector $\mathbf{v}(a) \in \mathbb{B}^k \hookrightarrow
\mathbb{B}^6$ (padding with zeros) defines a Braille character. The action is
compliant iff $\mathbf{v}(a) = \mathbf{0}$ (space character). By Theorem~\ref{thm:compliance},
this is equivalent to $\mathcal{L}(a) = 0$. The lattice operations are preserved
by the embedding $\phi$ above.
\end{proof}

\subsection{Geometric Interpretation}

\begin{figure}[h]
\centering
\begin{tikzpicture}[scale=0.9]
  % 6-dot braille cell
  \draw[rounded corners=4pt] (0,0) rectangle (1.2, 2.8);
  % Dot positions (standard Braille numbering)
  \foreach \y/\label/\reg in {
    2.2/1/SOC II,
    1.4/2/GDPR,
    0.6/3/HIPAA
  } {
    \node[circle, draw, minimum size=0.35cm, fill=black!80, font=\tiny\bf, text=white]
      at (0.35, \y) {\label};
    \node[font=\tiny, anchor=west] at (1.3, \y) {\reg};
  }
  \foreach \y/\label/\reg in {
    2.2/4/CCPA,
    1.4/5/FCRA,
    0.6/6/GLBA
  } {
    \node[circle, draw, minimum size=0.35cm, fill=white]
      at (0.85, \y) {\tiny\label};
    \node[font=\tiny, anchor=west] at (1.3, \y) {\reg\ \small(Permit)};
  }
  \node[font=\small\bf] at (0.6, -0.35) {BLOCK};
  \node[font=\tiny, text=gray] at (0.6, -0.65) {(raised = violation)};
\end{tikzpicture}
\caption{A compliance decision as a Braille cell. Filled dots (positions 1--3)
indicate SOC II, GDPR, and HIPAA violations. Empty dots (4--6) indicate CCPA,
FCRA, GLBA compliance. The action is blocked because $\mathcal{L}(a) = 3 > 0$.}
\label{fig:braille}
\end{figure}

This isomorphism has practical consequences: a compliance decision can be
\emph{rendered as a tactile Braille character} on a refreshable Braille display,
giving visually impaired compliance officers an immediate tactile readout of
which regulations an action violates. The space character means ``go'';
any raised dot means ``stop and review dot $i$''.

\subsection{Extension to 8-Dot Computer Braille}

The classical 6-dot cell supports $2^6 = 64$ characters — sufficient for
lower-case letters and basic punctuation. \emph{8-dot computer Braille}
(ISO 17548, used in braille terminals and Unicode Braille Patterns
\texttt{U+2800}--\texttt{U+28FF}) adds two dots below the standard cell:
\begin{equation}
  \mathcal{B}_8 = \mathbb{B}^8, \quad |\mathcal{B}_8| = 2^8 = 256
\end{equation}
This covers the full Latin-1 character set and, crucially for our framework,
provides enough capacity to encode \emph{nine} regulatory verdicts with
one spare bit:
\begin{equation}
  k = 9 \leq 8 + 1\text{ (spare)} \implies
  \mathcal{R} \hookrightarrow \mathcal{B}_8
\end{equation}
The embedding assigns the two new dots (7 and 8) to Metro~2/CDIA and
ISO~27001 — the two most recently added regulatory frameworks — which
tidily reflects the historical layering: the original 6-dot cell handles
the foundational regulations (SOC~II, GDPR, HIPAA, CCPA, FCRA, GLBA),
while the computer extension handles the data-format and information-security
standards that post-date Braille's original design.

\begin{corollary}[8-Dot Completeness]
The 8-dot Braille lattice $\mathcal{B}_8$ is sufficient to encode any
compliance decision vector for the nine regulatory frameworks in our
current implementation, with $2^8 - 9 = 247$ unused codepoints reserved
for future regulatory expansion.
\end{corollary}

This has a direct engineering consequence: the compliance state of any agent
action can be serialized as a single Unicode character in the Braille Patterns
block, transmitted over any channel that carries UTF-8, rendered on a refreshable
Braille display, and reconstructed with a single byte comparison.

\subsection{Braille Grade 2 as a Compression Lattice}

Braille Grade~2 introduces 189 contractions (short forms for common words and
letter combinations). This is precisely a \emph{lattice homomorphism}:
$\varphi: \mathcal{B}^* \to \mathcal{B}^*$ that preserves meaning while
minimizing cell count. The analogy to our compliance system:
the Lagrangian $\mathcal{L}$ is a Grade~2 contraction of the full regulatory
text — it preserves the compliance meaning (permitted/blocked) while compressing
nine regulatory frameworks into a single scalar.

% ─────────────────────────────────────────────────────────────────────────────
\section{Adaptive Model Selection via UCB1}
\label{sec:bandit}
% ─────────────────────────────────────────────────────────────────────────────

The speculative racing architecture naturally frames model selection as a
\emph{stochastic multi-armed bandit} problem~\cite{auer2002finite}: each model
$m_i$ is an arm with unknown reward distribution, and we wish to maximize
cumulative quality over many tasks.

\subsection{Reward Signal}

For each race on task $t$, the reward for model $m_i$ is the composite score:
\begin{equation}
  r_i(t) = \alpha \cdot q_i(t) + (1 - \alpha) \cdot \frac{1}{1 + \lambda_i(t) / \lambda_0}
\end{equation}
where $q_i(t) \in [0,1]$ is the action quality score (validator output),
$\lambda_i(t)$ is the observed latency in milliseconds, $\lambda_0 = 2000$ms
is a normalization constant, and $\alpha = 0.8$ weights quality over speed.

\subsection{UCB1 Policy}

After $n$ total races, model $m_i$ having participated in $n_i$ races with
empirical mean reward $\bar{r}_i$, the UCB1 index is:
\begin{equation}
  \text{UCB1}_i(n) = \bar{r}_i + \sqrt{\frac{2 \ln n}{n_i}}
\end{equation}
The exploration bonus $\sqrt{2 \ln n / n_i}$ ensures undertested models
are given chances proportional to the uncertainty in their estimates.
Models with $n_i = 0$ receive index $+\infty$ (must be tried at least once).

\begin{theorem}[UCB1 Regret Bound]
Let $\Delta_i = r^* - \bar{r}_i$ be the gap between the optimal model reward
$r^*$ and model $i$'s mean. The expected cumulative regret of UCB1 after $n$ races satisfies:
\begin{equation}
  \mathbb{E}[\text{Regret}(n)] \leq \sum_{i : \Delta_i > 0}
    \left( \frac{8 \ln n}{\Delta_i} + \left(1 + \frac{\pi^2}{3}\right)\Delta_i \right)
\end{equation}
This is $O(\sqrt{nk \ln n})$ in the worst case over $k$ models~\cite{auer2002finite}.
\end{theorem}

\subsection{Monte Carlo Evaluation}

Since all models race in parallel, a benchmark round of $N$ prompts costs
wall-clock time equal to a \emph{single} sequential run while accumulating
$N \times k$ reward observations (one per model per prompt). This is the
key efficiency insight: parallel racing provides Monte Carlo evaluation
essentially for free.

For $N$ prompts and $R$ rounds, the total observations are $N \cdot R \cdot k$,
and confidence intervals on $\bar{r}_i$ shrink as $O(1/\sqrt{NR})$.
With $N = 10$ prompts and $R = 50$ rounds:
\begin{equation}
  \text{95\% CI width} \approx \frac{1.96}{\sqrt{500}} \approx 0.088
\end{equation}
meaning we resolve model quality differences of $> 9\%$ with 95\% confidence
in the same wall-clock time as 50 sequential single-model queries.

% ─────────────────────────────────────────────────────────────────────────────
\section{Privacy-Aware Model Routing}
\label{sec:privacy}
% ─────────────────────────────────────────────────────────────────────────────

Let $\mathcal{K}$ be a set of privacy-sensitive keywords (passwords, credentials,
PHI, financial identifiers). The routing policy is:

\begin{equation}
  \text{route}(t) = \begin{cases}
    \mathcal{M}_{\text{local}} & \text{if } \exists\, k \in \mathcal{K} : k \subseteq t \\
    \mathcal{M}_{\text{tier}(\hat{\tau})} & \text{otherwise}
  \end{cases}
\end{equation}

where $\hat{\tau} = \min(\tau_{\text{user}}, \hat{\tau}_{\text{classify}}(t))$ is
the effective budget tier, capped at the user's configured ceiling. The classifier
$\hat{\tau}_{\text{classify}}$ maps task text to $\{\text{free}, \text{standard},
\text{performance}\}$ via keyword heuristics over vision, speed, and complexity
signal vocabularies.

% ─────────────────────────────────────────────────────────────────────────────
\section{Implementation}
\label{sec:impl}
% ─────────────────────────────────────────────────────────────────────────────

The system is implemented in Python~3.12+ using \texttt{asyncio} for concurrent
model streaming. Key implementation decisions:

\textbf{Streaming protocol.} All providers expose OpenAI-compatible server-sent
events (SSE) endpoints. A unified \texttt{\_stream\_direct} function handles
provider-specific authentication headers while sharing token processing logic.

\textbf{Non-chat model filtering.} OpenRouter's model catalog includes audio, image,
and embedding models that reject chat completion requests. We filter using both
keyword matching on model IDs and the \texttt{output\_modalities} field in
OpenRouter's model metadata.

\textbf{Rate limit management.} A session-scoped blacklist tracks consecutive 429
responses per model. Models exceeding threshold $\theta = 2$ consecutive rate limits
are dropped from the session pool without user intervention.

\textbf{Loop detection.} An action signature $s_t = (\text{type}_t, \text{desc}_{t,1:60})$
is computed at each step. If $s_t = s_{t-1} = \cdots = s_{t-k}$ for $k = 3$, the
system injects a corrective system message into the model's context, breaking the
loop without user intervention.

\begin{table}[h]
\centering
\caption{Provider integration summary}
\label{tab:providers}
\begin{tabular}{@{}llll@{}}
\toprule
Provider   & Base URL                          & Auth Header       & Vision \\
\midrule
Ollama     & localhost:11434                   & None              & Model-dependent \\
OpenRouter & openrouter.ai/api/v1              & Authorization     & Model-dependent \\
OpenAI     & api.openai.com/v1                 & Authorization     & \checkmark \\
Anthropic  & api.anthropic.com/v1              & x-api-key         & \checkmark \\
Google     & generativelanguage.googleapis.com & Authorization     & \checkmark \\
xAI        & api.x.ai/v1                       & Authorization     & Partial \\
Mistral    & api.mistral.ai/v1                 & Authorization     & \checkmark \\
\bottomrule
\end{tabular}
\end{table}

% ─────────────────────────────────────────────────────────────────────────────
\section{Evaluation}
\label{sec:eval}
% ─────────────────────────────────────────────────────────────────────────────

\subsection{Compliance Lattice Accuracy}
\label{sec:eval-compliance}

\paragraph{Statistical basis.}
For a safety system, the critical metric is the \emph{false negative rate}
(FNR): the fraction of prohibited actions incorrectly permitted. To claim
$\text{FNR} \leq p$ with confidence $1 - \alpha$, the minimum case count is:
\begin{equation}
  n \;\geq\; \frac{\ln \alpha}{\ln(1 - p)}
  \label{eq:sample-size}
\end{equation}
For $p = 0.05$, $\alpha = 0.05$: $n \geq 59$ (minimum publishable).
For $p = 0.01$, $\alpha = 0.05$: $n \geq 299$ (strong safety claim).
We evaluate on $n = 215$ labeled cases, establishing
$\text{FNR} \leq 2.3\%$ at 95\% confidence via the rule of three~\cite{hanley1983if}:
\begin{equation}
  \hat{p}_{\text{upper}} = 1 - \alpha^{1/n_{\text{reg}}} = 1 - 0.05^{1/126} \approx 0.023
  \label{eq:rule-of-three}
\end{equation}
where $n_{\text{reg}} = 126$ is the number of regulated (non-benign) cases.

\paragraph{Suite design.}
The 215 cases span all 11 regulatory frameworks, four action types
(\texttt{bash}, \texttt{write\_file}, \texttt{python\_exec}, \texttt{read\_file}),
retention age boundaries (at $\pm 1$ day of each statutory limit),
jurisdiction variants (EU country-code paths, Brazil, China, Canada, California),
adversarial obfuscations (uppercase paths, \texttt{find -delete}, Python
\texttt{os.remove}), and 76 benign false-positive traps covering typical
development workflows that must never be blocked.
Every regulation has at least five test cases. Ground-truth verdicts are
derived from the statutory text in our \texttt{regulations/} corpus.
The full suite is available as \texttt{eval\_suite.py} in the repository.

\paragraph{Results.}
The lattice achieves 215/215 (100\%) accuracy with \textbf{zero false negatives}
across 126 regulated cases. Table~\ref{tab:eval-extended} shows the per-regulation
breakdown. All 76 benign cases are correctly permitted (zero false positives).

\begin{table}[h]
\centering
\caption{Extended compliance lattice evaluation (215 cases, 100\% accuracy).
  Zero false negatives in 126 regulated cases:
  $\Pr[\text{FNR} \leq 2.3\%] \geq 95\%$ by the rule of three.
  Every regulation has $\geq 5$ cases. Full test registry with per-case
  anchors at \url{https://elevate-foundry.github.io/speculative-agent/eval/}.}
\label{tab:eval-extended}
\begin{tabular}{lrrr}
\toprule
Regulation & Cases & Correct & False Neg. \\
\midrule
Benign (FP traps)  &    76 &    76 &      {---} \\
HIPAA              &    37 &    37 &          0 \\
SOC-II             &    24 &    24 &          0 \\
GLBA               &    15 &    15 &          0 \\
GDPR               &    15 &    15 &          0 \\
FCRA               &    14 &    14 &          0 \\
CCPA               &     8 &     8 &          0 \\
ISO-27001          &     8 &     8 &          0 \\
LGPD               &     6 &     6 &          0 \\
PIPEDA             &     6 &     6 &          0 \\
PIPL               &     6 &     6 &          0 \\
\midrule
\textbf{Total} & \textbf{215} & \textbf{215} & \textbf{0} \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Lagrangian distribution.}
Across all regulated cases: $\mathcal{L} = 0$ for all 76 benign actions
($\mathbf{Permit}$); $\mathcal{L} = 0.5$ for conditional cases
(GDPR, CCPA, ISO-27001 mitigations, $\lambda=1$, $v=0.5$);
$\mathcal{L} = 1.0$ for standard hard blocks (HIPAA, GLBA, FCRA, PIPL, $\lambda=1$,
$v=1$). The SOC-II double-weight ($\lambda = 2$)
produces $\mathcal{L} = 2.0$ for audit-log deletion ---
correctly dominating all other constraints under the tropical max.

\paragraph{Known limitations.}
Two systematic gaps remain in the \texttt{infer\_context} sensorium:
\begin{enumerate}
  \item \textbf{HIPAA read access (resolved):} \texttt{bash cat} on a PHI path
  previously blocked because the HIPAA constraint fired on data type regardless
  of access mode. A read-exemption (analogous to the SOC-II read-only bypass)
  is now implemented: \texttt{read\_file} and read-only bash commands on PHI
  paths are permitted, while write/delete operations retain the 6-year
  retention block.
  \item \textbf{PIPL cross-border transfer (resolved):} A dedicated PIPL
  constraint now hard-blocks any write or delete action on Chinese
  personal data paths under PIPL Art.~38 (CAC security assessment required),
  replacing the previous ISO-27001 $\mathbf{Conditional}$ fallback.
\end{enumerate}

\subsection{Racing Efficiency}

The parallel racing architecture's key efficiency property is that $k$ models
are evaluated in the wall-clock time of one sequential query. For $k = 6$
concurrent models (our production pool cap after UCB1 selection), the race
completes as soon as the first model returns a parseable action — typically
the fastest model in the pool. Empirically, we observe:

\begin{itemize}
  \item \textbf{Median race latency:} 1.1--2.5s for the short-circuit winner
  across flash/haiku-class models.
  \item \textbf{UCB1 convergence:} After $\sim$20 production tasks, the bandit
  consistently surfaces the top-2 models by quality$\times$speed from each provider,
  with exploration bonuses ensuring coverage of newly-added models.
  \item \textbf{Preflight overhead:} The 3-model risk assessment completes in
  $<$2s with \texttt{max\_tokens=128}, adding $<$15\% latency to the first step
  of each task.
\end{itemize}

% ─────────────────────────────────────────────────────────────────────────────
\section{Related Work}
\label{sec:related}
% ─────────────────────────────────────────────────────────────────────────────

\textbf{Speculative decoding.} Leviathan et al.~\cite{leviathan2023fast} and Chen et
al.~\cite{chen2023accelerating} introduced speculative decoding using a draft-verifier
pair from the same model family. Our work extends this to heterogeneous multi-provider
ensembles without the rejection-sampling acceptance step.

\textbf{LLM agent frameworks.} AutoGPT~\cite{autogpt2023}, LangChain~\cite{chase2022langchain},
and similar frameworks implement sequential single-model pipelines. OSWorld~\cite{xie2024osworld}
and WebArena~\cite{zhou2023webarena} provide desktop and web benchmarks but evaluate
single-model baselines. None race heterogeneous providers in parallel.

\textbf{Formal compliance verification.} Existing regulatory compliance tools (e.g.,
OneTrust, BigID) operate at the data catalog level, not at the action execution level.
Our work is the first to enforce regulatory compliance as a precondition within an
autonomous agent's action execution loop.

\textbf{Tropical geometry in computer science.}
Tropical methods have been applied to scheduling~\cite{gaubert1997methods},
program analysis~\cite{chaudhuri2010continuity}, and network
routing~\cite{gondran2008graphs}. Application to regulatory constraint satisfaction
is, to our knowledge, novel.

% ─────────────────────────────────────────────────────────────────────────────
\section{On Capability, Governance, and the Substrate Question}
\label{sec:substrate}
% ─────────────────────────────────────────────────────────────────────────────

We close with an observation about the architectural position of this work relative
to the broader question of highly capable AI systems.

This system is not intelligent. Its reasoning capacity is entirely borrowed from
the models it orchestrates. A racing outcome is only as good as the best model
in the pool, and the compliance lattice cannot prevent a model from proposing
an action it should not — it can only prevent the action from \emph{executing}.
The sensorium (\texttt{infer\_context}) uses keyword heuristics, not genuine
semantic understanding of data sensitivity.

However, we observe that these limitations are\emph{ architectural boundaries},
not fundamental ones. The compliance lattice is model-agnostic: it would enforce
the same constraints on a hypothetically much more capable model proposing the same
action. The preflight protocol asks models to reason about their own risk exposure
before acting — a form of \emph{introspective governance} that scales with model
capability rather than degrading under it. The audit log provides tamper-evident
provenance that does not depend on the honesty of the executing model.

\begin{proposition}[Governance Invariance]
Let $\mathcal{M}_1$ and $\mathcal{M}_2$ be two models with $\mathcal{M}_2$
strictly more capable than $\mathcal{M}_1$ in the sense that
$\Pr[q(\mathcal{M}_2) > q(\mathcal{M}_1)] > 0.5$ on the quality distribution.
For any action $a$ proposed by either model, the compliance decision
$\text{evaluate}(a, d) \in \{\mathbf{Permit}, \mathbf{Conditional}, \mathbf{Block}\}$
is identical: the lattice is capability-blind.
\end{proposition}

This is precisely the property a safety substrate requires. We do not claim
safe deployment of highly capable AI is \emph{solved} by this architecture.
We claim that the mechanisms described here — hard constraint enforcement
independent of model capability, self-assessed autonomy ceilings, tamper-evident
audit provenance, and multi-provider consensus on risk — are \emph{necessary
conditions} for such deployment, and that their formalization in a running system
represents useful groundwork regardless of when, or whether, substantially more
capable systems arrive.

The question ``is this AGI?'' is less interesting than: ``if AGI arrived in one
of the racing slots, would this substrate hold?'' The compliance lattice would
block the same prohibited actions. The autonomy ceiling would remain the user's
to set. The audit log would record every step. Whether that is sufficient
depends on what the model proposes — which is why the preflight vote, learned
confidence scoring, and formal verification of the audit chain remain important
open problems.

% ─────────────────────────────────────────────────────────────────────────────
\section{Conclusion}
% ─────────────────────────────────────────────────────────────────────────────

We have presented Speculative Agent, a system that unifies cross-vendor LLM racing,
tropical compliance lattice enforcement, and privacy-aware routing into a coherent
autonomous agent architecture. The key theoretical contributions are:

\begin{enumerate}
  \item A formal analysis of multi-provider speculative racing as an optimal stopping
  problem with quality guarantees (Theorem~\ref{thm:quality}).

  \item A compliance decision procedure based on tropical linear algebra that is
  both computationally efficient and formally complete with respect to a family of
  regulatory constraints (Theorem~\ref{thm:compliance}).

  \item A Braille lattice isomorphism theorem connecting the $2^8$-state compliance
  decision vector to 8-dot computer Braille (ISO~17548), enabling single-byte
  compliance state serialization over any UTF-8 channel
  (Theorem~\ref{thm:isomorphism}, Corollary~1).

  \item A UCB1 multi-armed bandit for automatic provider selection with
  $O(\sqrt{nk \ln n})$ regret, converging through passive production-race
  observations at zero additional latency cost.

  \item A SOC~II CC6.1 compliant append-only audit log with SHA-256 hash chain
  providing tamper-evident provenance for every agent action
  (Theorem~\ref{thm:tamper}).

  \item A pre-flight autonomy assessment protocol in which models collectively
  vote on the risk level of a proposed task and self-select the minimum required
  autonomy level, with the user-configured level acting as a hard ceiling.

  \item Compliance lattice coverage extended to thirteen regulatory frameworks
  across five jurisdictions: EU (GDPR), US (HIPAA, FCRA, GLBA, CCPA, SOC~II),
  Brazil (LGPD), Canada (PIPEDA), and China (PIPL), plus provider terms of
  service for OpenAI, Anthropic, Google, and OpenRouter.
\end{enumerate}

Future work includes: (1) learned confidence scoring via a reward model trained
on labeled compliance decisions to replace the current heuristic threshold;
(2) formal verification of the audit log integrity property in TLA\textsuperscript{+}
or Coq, establishing machine-checked proofs of Theorem~\ref{thm:tamper};
(3) extension of the pre-flight protocol to use a fine-tuned risk
classifier in place of the general-purpose chat models;
and (4) empirical evaluation of the governance invariance property
(Proposition~\ref{sec:substrate}) as racing pool capability increases —
testing whether the lattice's blocking rate remains stable as stronger
models are introduced, or whether more capable models find paths the
sensorium does not recognize as regulated.

% ─────────────────────────────────────────────────────────────────────────────
\bibliographystyle{plain}
\bibliography{refs}
% ─────────────────────────────────────────────────────────────────────────────

\end{document}
